### 2018-04-21 assets/src/paper-stochastisch/main.tex
Add boilerplate paper + Makefile

\documentclass[natbib=true]{acmart}
\usepackage[utf8]{inputenc}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{sidebyside}
\usepackage{cleveref}
\usepackage{placeins}
\usepackage{colortbl}
\usepackage{bm}

\setcitestyle{numbers,sort&compress}

\title{An inquiry into the influence of parallelism and the choice of random
engines on the runtime and results of Stride}

\author{Niels Aerens}
\author{Thomas Avé}
\author{Tobia De Koninck}
\author{Robin Jadoul}

% \date{March 2018}

\begin{abstract}
\end{abstract}

\begin{document}

\maketitle

\section{Introduction}

\pagebreak

\section{Methods}

\section{Results}

\section{Future work}

\section{Postscriptum}

\clearpage

\bibliographystyle{ACM-Reference-Format}
\bibliography{references.bib}

\end{document}