### 2018-06-19 doc/latex/04-Code-quality.tex
Extra file!

%%!TEX root = ./UserManual.tex
\chapter{Code quality}
\label{chap:code-quality}

To ensure good code quality all compiler warnings were fixed our in some rare cases suppressed. The following compilers should give no warnings:
\begin{compactitem}
    \item GCC 6.4.0 on Linux
    \item GCC 7.3.0 on Linux
    \item GCC 7.3.1 on Linux
    \item GCC 8.1.1 on Linux
    \item Clang 5.0.0 on Linux
    \item Clang 5.0.1 on Linux
    \item Clang 5.0.2 on Linux
    \item Clang 6.0.0 on Linux
    \item AppleClang 8.1.0 on Darwin-16.7.0
    \item AppleClang 9.0.0 on Darwin-16.7.0
    \item AppleClang 9.1.0 on Darwin-17.4.0
\end{compactitem}

Clang-format was used to ensure a consistent code-style throughout the project. The used clang-format is 5.0.1.

The GUI's were manually tested on a macOS High Sierra system. We noticed that on one system, out of our control, an extra titlebar appeared. We were unable to reproduce this.
The used Qt version is